### 2020-04-20 paper.tex
Update tex file.

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{listings}
% Define slightly more reasonable Listings defaults
\lstset{
    basicstyle=\ttfamily\small,
    breaklines=true,
    prebreak=\raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
    frame=lines,
    showtabs=false,
    showspaces=false,
    showstringspaces=false,
    keywordstyle=\color[gray]{0.4}\bfseries,
    commentstyle=\color[gray]{0.65}\itshape,
    numbers=left,
    captionpos=b,
}

% Default fixed font does not support bold face
\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{8.5} % for bold
\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{8.5}  % for normal

% Custom colors
\usepackage{color}
\definecolor{deepblue}{rgb}{0,0,0.5}
\definecolor{deepred}{rgb}{0.6,0,0}
\definecolor{deepgreen}{rgb}{0,0.5,0}

% Python style for highlighting
\newcommand\pythonstyle{\lstset{
    language=Python,
    basicstyle=\ttm,
    aboveskip=10pt,
    belowskip=10pt,
    otherkeywords={self},             % Add keywords here
    keywordstyle=\ttb\color{deepblue},
    emph={MyClass,__init__},          % Custom highlighting
    emphstyle=\ttb\color{deepred},    % Custom highlighting style
    stringstyle=\color{deepgreen},
    numbers=none,
    frame=tb,                         % Any extra options here
    showstringspaces=false            % 
}}

% Python environment
\lstnewenvironment{python}[1][]
{
    \pythonstyle
    \lstset{#1}
}
{}

\usepackage{color}
\usepackage{fancyvrb}
\usepackage{fvextra}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, breaklines=true}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}

\usepackage[unicode=true]{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Devito for large scale elastic modelling and anisotropic inversion.},
            colorlinks=true,
            citecolor=black,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{longtable,booktabs}
\usepackage{caption}
\usepackage{subfig}
\captionsetup[subfloat]{margin=1em}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\bibliographystyle{IEEEtran}

\begin{document}

\title{Devito for large scale elastic modelling and anisotropic inversion}

\author{
\IEEEauthorblockN{1\textsuperscript{st} Louboutin Mathias}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}\\
Antlanta, GA \\
mlouboutin3@gatech.edu}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Fabio Luporini}
\IEEEauthorblockA{\textit{Devito Codes}}
\and
\IEEEauthorblockN{3\textsuperscript{th} Rhodri Nelson}
\IEEEauthorblockA{\textit{Imperial College London}}
\and
\IEEEauthorblockN{4\textsuperscript{rd} Philipp Witte}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}}
\and
\IEEEauthorblockN{5\textsuperscript{th} George Bisbas}
\IEEEauthorblockA{\textit{Imperial College London}}
\and
\IEEEauthorblockN{6\textsuperscript{th} Jan Thorbecke}
\IEEEauthorblockA{\textit{TU-Delft}}
\and
\IEEEauthorblockN{7\textsuperscript{th} Felix J. Herrmann}
\IEEEauthorblockA{\textit{Georgia Institute of Technology}}
\and
\IEEEauthorblockN{8\textsuperscript{th} Gerard J. Gorman}
\IEEEauthorblockA{\textit{Imperial College London}}
}

\maketitle

\begin{abstract}\label{abstract}

\href{https://github.com/devitocodes/devito}{Devito} is an open-source
Python project based on domain-specific language and compiler
technology. Driven by the requirements of rapid HPC applications
development in exploration seismology, the language and compiler have
evolved significantly since inception. Sophisticated boundary
conditions, tensor contractions, sparse operations and features such as
staggered grids and sub-domains are all supported; operators of
essentially arbitrary complexity can be generated. To accommodate this
flexibility whilst ensuring performance, data dependency analysis is
utilized to schedule loops and detect computational-properties such as
parallelism. In this article, the generation and simulation of
MPI-parallel propagators (along with their adjoints) for the
pseudo-acoustic wave-equation in tilted transverse isotropic media and
the elastic wave-equation are presented. Simulations are carried out on
industry scale synthetic models in a HPC Cloud system and reach a
performance of 28TFLOP/s, hence demonstrating Devito's suitability for
production-grade seismic inversion problems.
\end{abstract}

\section{Introduction}\label{introduction}

Seismic imaging methods such Reverse Time Migration (RTM) and
Full-waveform inversion (FWI) rely on the numerical solution of an
underlying system of partial differential equations (PDEs), most
commonly some manifestation of the wave-equation. In the context of FWI,
the finite-difference (FDM) and the spectral-element (SEM) methods are
most frequently used to solve the wave-equation, with FDM methods
dominating within the seismic exploration community \cite{lyu2020}.
Various forms of the wave-equation and modelling strategies for FWI are
detailed in \cite{fichtner2011}.

Despite the theory of FWI dating back to the 1980s
\cite{tarantola}, among the first successful
expositions on real 3D data was presented in \cite{sirgue2009}. Other
studies utilizing FDM within the FWI workflow include
\cite{ratcliffe2011, petersson2013}. The aforementioned studies
approximate the underlying physics via the acoustic wave-equation;
higher fidelity models solving the non-isotropic elastic wave-equation
have been developed in, e.g.,
\cite{osti_1468379, osti_1561580, osti_1561581, sava1, sava2}. Owing to
the flexibility of the mathematical discretizations that can be
utilized, along with the ability to describe problems on complex meshes,
there has also been a great deal of interest in utilizing SEM to solve
inversion problems \cite{peter2011, krebsdg}. The recent study
\cite{trinh2019} presents an efficient SEM based inversion scheme using
a viscoelastic formulation of the wave-equation.

It is generally accepted that the PDE solver utilized within an
inversion workflow must satisfy the following criteria
\cite{virieuxmodelling}: - Efficient for multiple-source modeling - The
memory requirement of the modeling - The ability of a parallel algorithm
to use an increasing number of processors - Ability of the method to
process models of arbitrary levels of heterogeneity - Reduce the
nonlinearity of FWI - Feasibility of the extension of the modeling
approach to more realistic physical descriptions of the media.

It is with these specifications in mind that
\href{https://github.com/devitocodes/devito}{Devito}, a symbolic domain
specific language (DSL) and compiler for the automatic generation of
finite-difference stencils, has been developed. Originally deigned to
accelerate research and development in exploration geophysics, the high-level
interface, previously described in detail in \cite{devito-api}, is built on top
of \texttt{SymPy} \cite{sympy} and is inspired by the underlying mathematical
formulations and other DSLs such as FEniCS \cite{fenics} and Firedrake
\cite{firedrake}. This interface allows the user to formulate wave-equations,
and more generally time-dependent PDEs in a simple and mathematically coherent
way. The Devito compiler then automatically generates finite-difference
stencils from these mathematical expressions. One of the main advantages of
\href{https://github.com/devitocodes/devito}{Devito} over other
finite-difference DSLs is that generic expressions such as sparse operations
(i.e.~point source injection or localized measurements) are fully supported and
expressible in a high-level fashion. The second component of
\href{https://github.com/devitocodes/devito}{Devito} is its compiler (c.f
\cite{devito-compiler}) that generates highly optimized C code. The generated
code is then compiled at runtime for the hardware at hand.

Previous work focused on the DSL and compiler to highlight the potential
application and use cases of Devito. Here, we present a series of
extensions and applications to large-scale three-dimensional problem
sizes as encountered in exploration geophysics. These experiments are
carried out in Cloud-based HPC systems and include elastic forward
modelling using distributed-memory parallelism and imaging based on the
tilted transverse isotropic (TTI) wave-equation
(\cite{virieux, thomsen1986, zhang-tti, duveneck, louboutin2018segeow}).
These proof of concepts highlight two critical features: first, the
ability of the symbolic interface and the compiler to translate to
large-scale adjoint-based inversion problems that require massive
compute (since thousands of PDEs are solved) as well as large amounts of
memory (since the adjoint state method requires the forward model to be
saved in memory). Secondly, through the elastic modelling example, we
demonstrate that \href{https://github.com/devitocodes/devito}{Devito}
now fully supports and automates vectorial and second order tensorial
staggered-grid finite-differences with the same high-level API
previously presented for scalar fields defined on cartesian grids.

This paper is organized as follows: first, we provide a brief overview
of \href{https://github.com/devitocodes/devito}{Devito} and its symbolic
API and present the distributed memory implementation that allows
large-scale modeling and inversion by means of domain decomposition. We
then provide a brief comparison with a state of the art hand-coded wave
propagator to validate the performance previously benchmarked with the
roofline model
(\cite{patterson, devito-compiler, devito-api, louboutin2016ppf}).
Before concluding, results from the Cloud-based experiments discussed
above are presented, highlighting the vectorial and tensorial
capabilities of Devito.

\section{Overview of Devito}\label{overview-of-devito}

Devito \cite{devito-api} provides a functional language built on top of
\texttt{SymPy} \cite{sympy} to symbolically express PDEs at a
mathematical level and implements automatic discretization with the
finite-difference method. The language is by design flexible enough to
express other types of non finite-diffrence operators, such as
interpolation and tensor contractions, that are inherent to
measurement-based inverse problems. Several additional features are
supported, among which staggered grids, sub-domains, and stencils with
custom coefficients. The last major building block of a good PDE solver
are the boundary conditions, that for finite-difference methods are
notoriously various and often complicated. The system is, however,
sufficiently general to express them through composition of core
mechanisms. For example, free surface and perfectly-matched layers
(PMLs) boundary conditions can be expressed as equations -- just like
any other PDE equations -- over a suitable sub-domain.

It is up to the \href{https://github.com/devitocodes/devito}{Devito}
compiler to translate the symbolic specification into C code. The
lowering of the input language down to C consists of several compilation
passes, some of which introducing performance optimizations that are the
key to fast code. Next to classic stencil optimizations (e.g., cache
blocking, alignment, SIMD and OpenMP parallelism),
\href{https://github.com/devitocodes/devito}{Devito} applies a series of
FLOP-reducing transformations as well as aggressive loop fusion. For a
complete treatment, the interested reader should refer to
\cite{devito-compiler}.

\subsection{Symbolic language and
compilation}\label{symbolic-language-and-compilation}

In this section we illustrate the
\href{https://github.com/devitocodes/devito}{Devito} language showing
how to implement the acoustic wave-equation in isotropic media
%
\begin{equation}
\begin{cases}
 m \frac{d^2 u(t, x)}{dt^2} - \Delta u(t, x) = \delta(xs) q(t) \\
 u(0, .) = \frac{d u(t, x)}{dt}(0, .) = 0 \\
 d(t, .) = u(t, xr).
 \end{cases}
\label{acou}
\end{equation}
%
 The core of the \href{https://github.com/devitocodes/devito}{Devito}
symbolic API consists of three classes:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \texttt{Grid}, a representation of the discretized model.
\item
  \texttt{(Time)Function}, a representation of spatially (and time-)
  varying variables defined on a \texttt{Grid} object.
\item
  \texttt{Sparse(Time)Function} a represention of (time-varying) point
  objects on a \texttt{Grid} object, generally unaligned with respect to
  the grid points, hence called ``sparse''.
\end{itemize}

A \texttt{Grid} represents a discretized finite n-dimensional space and
is created as follows

\begin{python}
from devito import Grid
grid = Grid(shape=(nx, ny, nz),
            extent=(ext_x, ext_y, ext_z),
            origin=(o_x, o_y, o_z))
\end{python}

where \texttt{(nx,\phantom{\ }ny,\phantom{\ }nz)} are the number of grid
points in each direction,
\texttt{(ext\_x,\phantom{\ }ext\_y,\phantom{\ }ext\_z)} is the physical
extent of the domain in physical units (i.e \texttt{m}) and
\texttt{(o\_x,\phantom{\ }o\_y,\phantom{\ }o\_z)} is the origin of the
domain in the same physical units. The object \texttt{grid} contains all
the information related to the discretization such as the grid spacing.
We use \texttt{grid} to create the symbolic objects that will be used to
express the wave-equation. First, we define a spatially varying model
parameter \texttt{m} and a time-space varying field \texttt{u}

\begin{python}
from devito import Function, TimeFunction
m = Function(name="m", grid=grid,
             space_order=so)
u = TimeFunction(name="u", grid=grid,
                 space_order=so,
                 time_order=to)
\end{python}

where \texttt{so} is the spatial discretization order and \texttt{to}
the time discretization order that is used for the generation of the
finite-difference stencil. Second, we define point-wise objects such as
point sources \texttt{src}, located at physical coordinates $x_s$, and
receiver (measurement) objects \texttt{d}, with sensors located at the
physical locations $x_r$

\begin{python}
from devito import SparseTimeFunction
s = SparseTimeFunction(name="src",
                       grid=grid, npoint=1,
                       coordinates=x_s)
d = SparseTimeFunction(name="d", grid=grid,
                       npoint=1, nt=nt,
                       coordinates=x_r)
\end{python}

The source term is handled separately from the PDE as a pointwise
operation called \texttt{injection}, while the measurement is handled
with an interpolation. By default,
\href{https://github.com/devitocodes/devito}{Devito} initializes all
\texttt{Function} data to 0, and thus automatically satisfies the zero
Dirichlet condition at \texttt{t=0}. The isotropic acoustic
wave-equation can then be implemented in
\href{https://github.com/devitocodes/devito}{Devito} as below

\begin{python}
from devito import solve, Eq, Operator
eq = m * u.dt2 - u.laplace
update = Eq(u.forward, solve(eq, u.forward))
src_eqns = s.inject(u.forward, expr=s*dt**2/m)
d_eqns = d.interpolate(u)
\end{python}

To trigger compilation one needs to pass the constructed equations to an
\texttt{Operator}.

\begin{python}
from devito import Operator
op = Operator(update + src_eqns + d_eqns)
\end{python}

The first compilation pass processes the equations individually. The
equations are lowered to an enriched representation, while the
finite-difference constructs (e.g., derivatives) are translated into
actual arithmetic operations. Subsequently, data dependency analysis is
used to compute a performance-optimized topological ordering of the
input equations (e.g., to maximize the likelihood of loop fusion) and to
group them into so called ``clusters''. Basically, a cluster will
eventually be a loop nest in the generated code, and consecutive
clusters may share some outer loops. The ordered sequence of clusters
undergoes several optimization passes, including cache blocking and
FLOP-reducing transformations. It is then further lowered into an
abstract syntax tree, and it is on such representation that parallelisms
is introduced (SIMD, shared-memory, MPI). Finally, all remaining
low-level aspects of code generation are handled, among which the most
relevant one is the data management (e.g., definition of variables,
transfers between host and device).

The output of the \href{https://github.com/devitocodes/devito}{Devito}
compiler for the running example used in this section is available at
\href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
in \texttt{acou-so8.c}.

\subsection{Distributed-memory
parallelism}\label{distributed-memory-parallelism}

We here provide a succinct description of distributed-memory parallelism
in Devito; the interested reader should refer to the MPI tutorial at
\href{https://github.com/devitocodes/devito/blob/master/examples/mpi/overview.ipynb}{mpi-notebook}
for thorough explanations and practical examples.

Devito implements distributed-memory parallelism on top of MPI. The
design is such that users can almost entirely abstract away from it and
reuse non-distributed code as is. Given any
\href{https://github.com/devitocodes/devito}{Devito} code, just running
it as

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DEVITO_MPI=}\DecValTok{1} \NormalTok{mpirun -n X python ...}
\end{Highlighting}
\end{Shaded}

triggers the generation of code with routines for halo exchanges. The
routines are scheduled at a suitable depth in the various loop nests
thanks to data dependency analysis. The following optimizations are
automatically applied:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  redundant halo exchanges are detected and dropped;
\item
  computation/communication overlap, with prodding of the asynchronous
  progress engine by a designated thread through repeated calls to
  \texttt{MPI\_Test};
\item
  a halo exchange is placed as far as possible from where it is needed
  to maximize computation/communication overlap;
\item
  data packing and unpacking is threaded.
\end{itemize}

The domain decomposition occurs in Python upon creation of a
\texttt{Grid} object. Exploiting the MPI Cartesian topology abstraction,
\href{https://github.com/devitocodes/devito}{Devito} logically splits a
grid based on the number of available MPI processes (users are given an
``escape hatch'' to override Devito's default decomposition strategy).
\texttt{Function} and \texttt{TimeFunction} objects inherit the
\texttt{Grid} decomposition. For \texttt{SparseFunction} objects the
approach is different. Since a \texttt{SparseFunction} represents a
sparse set of points,
\href{https://github.com/devitocodes/devito}{Devito} looks at the
physical coordinates of each point and, based on the \texttt{Grid}
decomposition, schedules the logical ownership to an MPI rank. If a
sparse point lies along the boundary of two or more MPI ranks, then it
is duplicated in each of these ranks to be accessible by all neighboring
processes. Eventually, a duplicated point may be redundantly computed by
multiple processes, but any redundant increments will be discarded.

When accessing or manipulating data in a
\href{https://github.com/devitocodes/devito}{Devito} code, users have
the illusion to be working with classic NumPy arrays, while underneath
they are actually distributed. All manner of NumPy indexing schemes
(basic, slicing, etc.) are supported. In the implementation, proper
global-to-local and local-to-global index conversion routines are used
to propagate a read/write access to the impacted subset of ranks. For
example, consider the array

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{A = [[ }\DecValTok{1}\NormalTok{,  }\DecValTok{2}\NormalTok{,  }\DecValTok{3}\NormalTok{,  }\DecValTok{4}\NormalTok{],}
     \NormalTok{[ }\DecValTok{5}\NormalTok{,  }\DecValTok{6}\NormalTok{,  }\DecValTok{7}\NormalTok{,  }\DecValTok{8}\NormalTok{],}
     \NormalTok{[ }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{11}\NormalTok{, }\DecValTok{12}\NormalTok{],}
     \NormalTok{[}\DecValTok{13}\NormalTok{, }\DecValTok{14}\NormalTok{, }\DecValTok{15}\NormalTok{, }\DecValTok{16}\NormalTok{]])}
\end{Highlighting}
\end{Shaded}

which is distributed across 4 ranks such that \texttt{rank\phantom{\ }0}
contains the elements reading
\texttt{1,\phantom{\ }2,\phantom{\ }5,\phantom{\ }6},
\texttt{rank\phantom{\ }1} the elements
\texttt{3,\phantom{\ }4,\phantom{\ }7,\phantom{\ }8},
\texttt{rank\phantom{\ }2} the elements
\texttt{9,\phantom{\ }10,\phantom{\ }13,\phantom{\ }14} and
\texttt{rank\phantom{\ }3} the elements
\texttt{11,\phantom{\ }12,\phantom{\ }15,\phantom{\ }16}. The slicing
operation \texttt{A{[}::-1,\phantom{\ }::-1{]}} will then return

\begin{Shaded}
\begin{Highlighting}[]
    \NormalTok{[[ }\DecValTok{16}\NormalTok{, }\DecValTok{15}\NormalTok{, }\DecValTok{14}\NormalTok{, }\DecValTok{13}\NormalTok{],}
     \NormalTok{[ }\DecValTok{12}\NormalTok{, }\DecValTok{11}\NormalTok{, }\DecValTok{10}\NormalTok{,  }\DecValTok{9}\NormalTok{],}
     \NormalTok{[  }\DecValTok{8}\NormalTok{,  }\DecValTok{7}\NormalTok{,  }\DecValTok{6}\NormalTok{,  }\DecValTok{5}\NormalTok{],}
     \NormalTok{[  }\DecValTok{4}\NormalTok{,  }\DecValTok{3}\NormalTok{,  }\DecValTok{2}\NormalTok{,  }\DecValTok{1}\NormalTok{]])}
\end{Highlighting}
\end{Shaded}

such that now \texttt{rank\phantom{\ }0} contains the elements
\texttt{16,\phantom{\ }15,\phantom{\ }12,\phantom{\ }11} and so forth.

Finally, we remark that while providing abstractions for distributed
data manipulation, \href{https://github.com/devitocodes/devito}{Devito}
does not support natively any mechanisms for parallel I/O. The
distributed NumPy arrays, however, provide a generic and flexible
infrastructure for the implementation of parallel I/O for any sort of
file format (e.g., see \cite{witte2018alf}).

\section{Industry-scale 3D seismic imaging in anisotropic
media}\label{industry-scale-3d-seismic-imaging-in-anisotropic-media}

One of the main applications of seismic finite-difference modeling in
exploration geophysics is reverse-time migration (RTM), a wave-equation based
seismic imaging technique. Real-world seismic imaging presents a number of
challenges that make applying this method to industry-scale problem sizes
difficult. Firstly, RTM requires an accurate representation of the underlying
physics via sophisticated wave-equations such as the tilted-transverse isotropic
(TTI) wave-equation, for which both forward and adjoint implementations must to
be provided. Secondly, wave-equations must be solved for a large number of
independent experiments, where each individual PDE solve is itself expensive in
terms of FLOPs and memory usage. For certain workloads, limited domain
decomposition, that balances the domain size and the number of independent
experiments, and checkpointing techniques must to be adopted. In the following
sections, we describe an industry-scale seismic imaging problem that poses all
the aforementioned challenges, its implementation with Devito, and the results
of an experiment carried out on the Azure Cloud using a synthetic data set.

\subsection{Anisotropic wave-equation}\label{anisotropic-wave-equation}

In our seismic imaging case study, we use an anisotropic representation of the
physics called Tilted Transverse Isotropic (TTI) modeling \cite{thomsen1986}. This
representation for wave motion is one of the most widely used in exploration
geophysics since it captures the leading order kinematics and dynamics of
acoustic wave motion in highly heterogeneous elastic media where the medium
properties vary more rapidly in the direction perpendicular to sedimentary
strata \cite{alkhalifah2000, baysal1983, bubetti2012, bubetti2014,
bubesatti2016, chu2011, duveneck, fletcher, fowlertti2010,
louboutin2018segeow, whitmore1983, witte2016segpve, xu2014, ,zhang2005,
zhang2011, zhan2013}. The TTI wave-equation is an acoustic, low dimensional (4
parameters, 2 wavefields) simplification of the 21 parameter and 12 wavefields
tensorial equations of motions \cite{hooke}. This simplified representation is
parametrized by the Thomsen parameters $\epsilon(x), \delta(x)$ that relate to
the global (propagation over many wavelengths) difference in propagation speed
in the vertical and horizontal directions, and the tilt and azimuth angles
$\theta(x), \phi(x)$ that define the rotation of the vertical and horizontal
axis around the cartesian directions. However, unlike the scalar isotropic
acoustic wave-equation itself, the TTI wave-equation is extremely
computationally costly to solve and it is also not self-adjoint as shown in
\cite{louboutin2018segeow}.

The main complexity of the TTI wave-equation is that the rotation of the
symmetry axis of the physics leads to rotated second-order finite-difference
stencils. In order to ensure numerical stability, these rotated
finite-difference operators are designed to be self-adjoint (c.f.
\cite{zhang2011, duveneck}). For example, we define the rotated second order
derivative with respect to $x$ as:
%
\begin{equation}
\begin{aligned}
  G_{\bar{x}\bar{x}} &= D_{\bar{x}}^T D_{\bar{x}} \\
  D_{\bar{x}} &= \cos(\mathbf{\theta})\cos(\mathbf{\phi})\frac{\mathrm{d}}{\mathrm{d}x} + \cos(\mathbf{\theta})\sin(\mathbf{\phi})\frac{\mathrm{d}}{\mathrm{d}y} - \sin(\mathbf{\theta})\frac{\mathrm{d}}{\mathrm{d}z}.
\end{aligned}
\label{rot}
\end{equation}
%
 We enable the simple expression of these complicated stencils in
\href{https://github.com/devitocodes/devito}{Devito} as
finite-difference shortcuts such as \texttt{u.dx} where \texttt{u}
is a \texttt{Function}. Such shortcuts are enabled not only for the
basic types but for generic composite expressions, for example
\texttt{(u\phantom{\ }+\phantom{\ }v.dx).dy}. As a consequence, the
rotated derivative defined in~\ref{rot} is implemented with
\href{https://github.com/devitocodes/devito}{Devito} in two lines as:

\begin{python}
dx_u = cos(theta) * cos(phi) * u.dx +
       cos(theta) * sin(phi) * u.dy -
       sin(theta) * u.dz
dxx_u = (cos(theta) * cos(phi) * dx_u).dx.T +
        (cos(theta) * sin(phi) * dx_u).dy.T -
        (sin(theta) * dx_u).dz.T
\end{python}

Note that while the adjoint of the finite-difference stencil is enabled via the
standard Python \texttt{.T} shortcut, the expression needs to be reordered by hand
since the tilt and azymuth angles are spatially dependent and require to be
inside the second pass of first-order derivative. We can see from these simple
two lines that the rotated stencil involves all second-order derivatives
(\texttt{.dx.dx}, \texttt{.dy.dy} and \texttt{.dz.dz}) and all second-order
cross-derivatives (\texttt{dx.dy}, \texttt{.dx.dz} and \texttt{.dy.dz}) which
leads to a denser stencil support and higher computational complexity
(c.f. \cite{louboutin2016ppf}). For illustrative purposes, the complete
generated code for tti modeling with and without MPI is made available
at \href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
in \texttt{tti-so8.c} and \texttt{tti-so8-mpi.c}.

Owing to the very high number of floating-point operations (FLOPs) needed per
grid point for the weighted rotated Laplacian, this anisotropic wave-equation is
extremely challenging to implement. As we show in Table Table~\ref{ttiFLOPs},
and previously analysed in \cite{louboutin2016ppf}, the computational cost with
high-order finite-difference is in the order of thousands of FLOPs per grid
point without optimizations. The version without FLOP-reducing optimizations is
a direct translation of the discretized operators into stencil expressions. The
version with optimizations employs transformations such as common
sub-expressions elimination, factorization, and cross-iteration redundancy
elimination -- the latter being key in removing redundancies introduced by mixed
derivatives. Implementing all of these techniques manually is inherently
difficult and laborious. Further, to obtain the desired performance improvements
it is necessary to orchestrate them with aggressive loop fusion (for data
locality), tiling (for data locality and tensor temporaries), and potentially
ad-hoc vectorization strategies (if rotating registers are used). While an
explanation of the optimization strategies employed by
\href{https://github.com/devitocodes/devito}{Devito} is beyond the scope of this
paper (see \cite{devito-compiler} for details), what is emphasized here is that
users can easily take full advantage of these optimization without needed to
concern themselves with the details.

\begin{table}
\centering
\begin{tabular}{ccc}
\toprule\addlinespace
spatial order & w/o optimizations & w/ optimizations\tabularnewline
\midrule
4 & 501 & 95\tabularnewline
8 & 539 & 102\tabularnewline
12 & 1613 & 160\tabularnewline
16 & 5489 & 276\tabularnewline
\bottomrule
\end{tabular}
\caption{Per-grid-point FLOPs of the finite-difference TTI wave-equation stencil
with different spatial discretization orders.}\label{ttiFLOPs}
\end{table}

It is evident that developing an appropriate solver for the TTI wave-equation,
an endeavor involving complicated physics, mathematics, and engineering, is
exceptionally time-consuming and can lead to thousands of lines of code even for
a single choice of discretization. Verification of the results is no less
complicated, any minor error is effectively untrackable and any change to the
finite-difference scheme or to the time-stepping algorithm is difficult to
achieve without substantial re-coding. Another complication stems from the fact
that practitioners of seismic inversion are often geoscientists, not computer
scientists/programmers. Low level implementations from non-specialists can often
often lead to poorly performing code. However, if research codes are passed to
specialists in the domain of low level code optimization they often lack the
necessary geophysical domain knowledge, resulting in code that may lack a key
feature required by the geoscientist. Neither situation is conducive to
addressing the complexities that come with implementing codes based on the
latest geophysical insights in tandem with those from high-performance
computing. With \href{https://github.com/devitocodes/devito}{Devito} on the
other hand, both the forward and adjoint equations can be implemented in a few
lines of Python code as illustrated with the rotated operator in \#rotxpy. The
low level optimization element of the development is then taken care of under
the hood by the \href{https://github.com/devitocodes/devito}{Devito} compiler.

The simulation of wave motion is only one aspect of solving problems in
seismology. During wave-equation based imaging, it is also required to compute
sensitivities (gradient) with respect to the quantities of interest. This
requirement imposes additional constraints on the design and implementation of
model codes as outlined in \cite{virieux}. Along with several factors, such as
fast setup time, we focused on correct and testable implementations for the
adjoint wave-equation and the gradient (action of the adjoint Jacobian)
\cite{louboutin2018segeow, louboutin2020THmfi}; exactness being a mandatory
requirement of gradient based iterative optimization algorithms.

\subsection{3D Imaging example on Azure}\label{d-imaging-example-on-azure}

We now demonstrate the scalability of
\href{https://github.com/devitocodes/devito}{Devito} to real-world applications
by imaging an industry-scale three-dimensional TTI subsurface model. This
imaging was carried out in the Cloud on Azure and takes advantage of recent work
to port conventional cluster code to the Cloud using a serverless approach. The
serverless implementation is detailed in \cite{witte2019TPDedas,
witte2019SEGedw} where the steps to run computationally and financially
efficient HPC workloads in the Cloud are described. This imaging project, in
collaboration with Azure, demonstrates the scalability and robustness of
\href{https://github.com/devitocodes/devito}{Devito} to large scale
wave-equation based inverse problems in combination with a cost-effective
serverless implementation of seismic imaging in the Cloud. In this example, we
imaged a synthetic three-dimensional anisotropic subsurface model that mimics a
realistic industry size problem with a realistic representation of the physics
(TTI). The physical size of the problem is \texttt{10kmx10kmx2.8km} discretized
on a \texttt{12.5m} grid with absorbing layers of width 40 grid points on each
side leading to \texttt{881x881x371} computational grid points ($\approx300$
Million grid points). The final image is the sum of 1500 single-source images:
100 single-source images were computed in parallel on the 200 nodes available
using two nodes per source experiment.

\textbf{\emph{Computational performance}}

We briefly describe the computational setup and the performance achieved for
this anisotropic imaging problem. Due to time constraints, and because the
resources we were given access to for this proof of concept with Azure were
somewhat limited, we did not have access to HPC virtual machines (VM) nor
Infiniband enabled ones. This experiment was carried out on
\texttt{Standard\_E64\_v3} and \texttt{Standard\_E64s\_v3} nodes which, while
not HPC VM nodes, are memory optimized thus allowing to the wavefield to be
stored in memory for imaging (TTI adjoint state gradient \cite{virieux,
louboutin2018segeow}). These VMs are Intel® Broadwell E5-2673 v4 2.3GH that are
dual socket, 32 physical cores (with hyperthreading enabled) and 432Gb of DRAM.
The overall inversion involved computing the image for 1500 source positions,
i.e.~solving 1500 forward and 1500 adjoint TTI wave-equations. A single image
required, in single precision, 600Gb of memory. Two VMs were used per source and
MPI set with one rank per socket (4 MPI ranks per source) and 100 sources were
imaged in parallel. The performance achieved was as follows:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  140 GFLOP/s per VM;
\item
  280 GFLOP/s per source;
\item
  28 TFLOP/s for all 100 running sources;
\item
  110min runtime per source (forward + adjoint + image computation).
\end{itemize}

We comment that if more resources were available, and because the imaging
problem is embarrassingly parallel over sources and can scale arbitrarily,the
imaging of all of the 1500 sources in parallel could have been attempted, which
theoretically leads to a performance of 0.4PFLOP/s.

\textbf{\emph{How performance was measured}}

The execution time is computed through Python-level timers prefixed by an MPI
barrier. The floating-point operations are counted once all of the symbolic
FLOP-reducing transformations have been performed during compilation.
\href{https://github.com/devitocodes/devito}{Devito} uses an in-house estimate
of cost, rather than \texttt{SymPy}'s estimate, to take care of some low-level
intricacies. For example, Devito's estimate ignores the cost of integer
arithmetic used for offset indexing into multi-dimensional arrays. To calculate
the total number of FLOPs performed,
\href{https://github.com/devitocodes/devito}{Devito} multiplies the
floating-point operations calculated at compile time by the size of the
iteration space, and it does that at the granularity of individual expressions.
Thanks to aggressive code motion, the amount of innermost-loop-invariant
sub-expressions in a \href{https://github.com/devitocodes/devito}{Devito}
Operator is typically negligible and therefore the
\href{https://github.com/devitocodes/devito}{Devito} estimate does not suffer
from this issue, or at least not, to the best of our knowledge, in a tangible
way. The Devito-reported GFLOP/s were also checked against those produced by
Intel Advisor on several single-node experiments: the differences -- typically
\href{https://github.com/devitocodes/devito}{Devito} underestimating the
achieved performance -- were always at most in the order of units, and therefore
negligible.

\textbf{\emph{Imaging result}}

The subsurface velocity model used in this study is an artificial anisotropic
model that is designed and built combining two broadly known and used
open-source SEG/EAGE acoustic velocity models that each come with realistic
geophysical imaging challenges such as sub-salt imaging. The anisotropy
parameters are derived from a smoothed version of the velocity while the tilt
angles were derived from a combination of the smooth velocity models and
vertical and horizontal edge detection. The final seismic image of the
subsurface model is displayed in Figure~\ref{OverTTI} and highlights the fact
that 3D seismic imaging based on a serverless approach and automatic code
generation is feasible and provides good results.

\begin{figure*}
\centering
\captionsetup[subfigure]{labelformat=empty}
\subfloat[]{\includegraphics[width=0.500\hsize]{./Figures/OverTTI1.png}}
\subfloat[]{\includegraphics[width=0.500\hsize]{./Figures/OverTTI2.png}}
\\
\subfloat[]{\includegraphics[width=0.500\hsize]{./Figures/OverTTI3.png}}
\subfloat[]{\includegraphics[width=0.500\hsize]{./Figures/OverTTI4.png}}
\caption{3D TTI imaging on a custom made model.}\label{OverTTI}
\end{figure*}

\cite{witte2019TPDedas} describe the serverless implementation of seismic
inverse problems in detail, including iterative algorithms for least-square
minimization problems (LSRTM). The 3D anisotropic imaging results were presented
as part of a keynote presentation at the EAGE HPC workshop in October 2019
\cite{herrmann2019EAGEHPCaii} and at the Rice O\&G HPC workshop
\cite{witte2019RHPCssi} in which the focus was on the serverless implementation
of seismic inversion algorithms in the Cloud. This work illustrates the
flexibility and portability of Devito: we were able to easily port a code
developed and tested on local hardware to the Cloud, with only minor
adjustments. Further, note that this experiment included the porting of
MPI-based code for domain decomposition developed on desktop computers to the
Cloud. Our experiments are reproducible using the instructions in a public
repository \href{https://github.com/slimgroup/Azure2019}{AzureTTI}, which
contains, among the other things, the Dockerfiles and Azure
\href{https://batch-shipyard.readthedocs.io}{batch-shipyard} setup. This example
can also be easily run on a traditional HPC cluster environment using, for
example, \href{https://github.com/slimgroup/JUDI.jl}{JUDI} or Dask~\cite{dask}
for parallelization over sources.

\section{Elastic modelling}\label{elastic-modelling}

While the subsurface image in the previous Section is obtained with
anisotropic propagators capable of mimicking the real physics, in order
to model both the kinematics and the amplitudes correctly elastic
propagators are required. These propagators are for example extremely
important for global seismology, as the Shear surface waves (component
ignored in TTI) are the most hazardous ones. In this section, we exploit
the tensor algebra language introduced in
\href{https://github.com/devitocodes/devito}{Devito} v4.0 to express
elastic waves in a compact and elegant notation.

The elastic isotropic wave-equation, parametrized by the Lamé parameters
$\lambda, \mu$ and the density $\rho$ reads:
%
\begin{equation}
\begin{aligned}
&\frac{1}{\rho}\frac{dv}{dt} = \nabla . \tau \\
&\frac{d \tau}{dt} = \lambda \mathrm{tr}(\nabla v) \mathbf{I}  + \mu (\nabla v + (\nabla v)^T)
\end{aligned}
\label{elas1}
\end{equation}
%
 where $v$ is a vector valued function with one component per cartesian
direction:
%
\begin{equation}
\begin{split}
v =  \begin{bmatrix} v_x(t, x, y) \\ v_y(t, x, y)) \end{bmatrix}
\end{split}
\label{partvel}
\end{equation}
%
 and the stress $\tau$ is a symmetric second-order tensor-valued
function:
%
\begin{equation}
\begin{aligned}
    \tau = \begin{bmatrix}\tau_{xx}(t, x, y) & \tau_{xy}(t, x, y)\\\tau_{xy}t, x, y) & \tau_{yy}(t, x, y)\end{bmatrix}.
\end{aligned}
\label{stress}
\end{equation}
%
 The discretization of such a set of coupled PDEs requires five
equations in two dimensions (two equations for the particle velocity and
three for stress) and nine equations in three dimensions (three particle
velocities and six stress equations). However, the mathematical
definition only requires two coupled vector/tensor-valued equations for
any number of dimensions.

\subsection{Tensor algebra language}\label{tensor-algebra-language}

We have augmented the
\href{https://github.com/devitocodes/devito}{Devito} language with
tensorial objects to enable a straightforward -- and at the same time
mathematically rigorous -- definition of high-dimensional PDEs, such as
the elastic wave-equation in Equation~\ref{elas1}. This effort was
inspired by \cite{ufl}, a functional language for finite element
methods.

The extended \href{https://github.com/devitocodes/devito}{Devito}
language introduces two new types, \texttt{VectorFunction} (and
\texttt{VectorTimeFunction}) for vectorial objects such as the particle
velocity, and \texttt{TensorFunction} (and \texttt{TensorTimeFunction})
for second-order tensor objects (matrices) such as the stress. These new
objects are constructed the exact same way as the scalar
\texttt{Function} objects. They also automatically implement staggered
grid and staggered finite-differences with the possibility of half-node
averaging. Each component of a tensorial object -- a (scalar)
\href{https://github.com/devitocodes/devito}{Devito} \texttt{Function}
-- is accessible via conventional vector notation (i.e.
\texttt{v{[}0{]},\phantom{\ }t{[}0,1{]},....}).

With this extended language, the elastic wave-equation can be defined in
only four lines as follows:

\begin{python}
from devito import VectorTimeFunction,
                   TensorTimeFunction
v = VectorTimeFunction(name="v", grid=grid,
                       space_order=so,
                       time_order=1)
tau = TensorTimeFunction(name="t", grid=grid,
                         space_order=so,
                         time_order=1)
u_v = Eq(v.forward,
         damp * (v + s/rho*div(tau)))
u_t = Eq(tau.forward,
         damp * (tau + s * (l * diag(div(v.forward)) + mu * (grad(v.forward) + grad(v.forward).T))))
\end{python}

The \texttt{SymPy} expressions created by these commands can be
displayed with \texttt{sympy.pprint} as shown in
Figure~\ref{PrettyElas}. This representation reflects perfectly the
mathematics while still providing computational portability and
efficiency through the
\href{https://github.com/devitocodes/devito}{Devito} compiler. The
complete generated code for the elastic wave-equation with and without
MPI is made available at
\href{https://github.com/mloubout/SC20Paper/tree/master/gencode}{CodeSample}
in \texttt{elastic-so8.c} and \texttt{elastic-so8-mpi.c}.

\begin{figure*}
\centering
\includegraphics[width=1.000\hsize]{./Figures/vel_symb.png}
\caption{Update stencil for the particle velocity. The stencil for
updating the stress component is left out for readability, as the
equation does not fit onto a single page. However, it can be found in
the \href{https://github.com/devitocodes/devito}{Devito} tutorial on
elastic modelling on github.}\label{PrettyElas}
\end{figure*}

\subsection{2D example}\label{d-example}

We show the elastic particle velocity and stress for a broadly
recognized 2D synthetic model, the elastic
Marmousi-ii\cite{versteeg927, marmouelas} model. The wavefields are
shown on Figure~\ref{ElasWf} and its corresponding elastic shot records
are displayed in Figure~\ref{ElasShot}. Thes two figures show that the
wavefield is as expected purely acoustic in the water layer
(\texttt{t\_\{xy\}=0}) and transition perfectly at the ocean bottom to
an elastic wavefield. We can also clearly see the Shear wave wavefront
in the subsurface (around 1km depth). Thes two plots show that we
effectively solve the elastic wave-equation with a high-level
mathematical definition of the PDE that abstracts away all the
computational technicalities such as staggered grid analysis. The shot
records in Figure~\ref{ElasShot} also match the original data generated
by the creator of the elastic model in \cite{marmouelas}.

\begin{figure*}
\centering
\includegraphics[width=1.000\hsize]{./Figures/marmou_snap.png}
\caption{Particle velocities and stress at time $t=3\text{s}$ for a
source at 10m depth and \texttt{x=5\textbackslash{}text\{km\}} in the
marmousi-ii model.}\label{ElasWf}
\end{figure*}

\begin{figure*}
\centering
\captionsetup[subfigure]{labelformat=empty}
\subfloat[]{\includegraphics[width=0.300\hsize]{./Figures/pressure_marmou.png}}
\subfloat[]{\includegraphics[width=0.300\hsize]{./Figures/vz_marmou.png}}
\subfloat[]{\includegraphics[width=0.300\hsize]{./Figures/vx_marmou.png}}
\caption{Seismic shot record for 5sec of modelling. \texttt{a} is the
pressure (trace of stress tensor) at the surface (5m depth), \texttt{b}
is the vertical particle velocity and \texttt{c} is the horizontal
particle velocity at the ocean bottom (450m depth).}\label{ElasShot}
\end{figure*}

\subsection{3D proof of concept}\label{d-proof-of-concept}

Finally, we modelled three dimensional elastic data in the Cloud to
demonstrate the scalability of
\href{https://github.com/devitocodes/devito}{Devito} to cluster-size
problems in the Cloud. The model we chose mimics the reference model in
geophysics known as the SEAM model \cite{fehler2011seam} that is a
three dimensional extreme-scale synthetic representation of the
subsurface. The physical dimension of the model are
\texttt{45kmx35kmx15km} then discretized with a grid spacing of
\texttt{20mx20mx10m} that led to a computational grid of
\texttt{2250x1750x1500} grid points (5.9 billion grid points). One of
the main challenges of elastic modelling is the extreme memory cost due
to the number of wavefield. For a three dimensional propagator, a
minimum of 21 fields need to be stored:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Three particle velocities with two time steps (\texttt{v.forward} and
  \texttt{v})
\item
  Six stress with two time steps (\texttt{tau.forward} and \texttt{tau})
\item
  Three model parameters \texttt{lamda}, \texttt{mu} and \texttt{rho}
\end{itemize}

These 21 fields, with the grid we just describe, leads to a minimum of
461Gb of memory for modelling only. For this experiment, we obtained
access to small HPC VM on azure called \texttt{Standard\_H16r} that are
16 cores Intel Xeon E5 2667 v3 with no hyperthreading and used 32 nodes
for a single source experiment (we solved a single wave-equation). We
used a 12th order discretization that leads to 2.8TFLOP/time-step to be
computed for this model and propagated the elastic wave for 16 seconds
(23000 time steps). The modelling ran in 16 hours which converts to
1.1TFLOP/s. While these number may seem low, the elastic kernel is
extremely memory bound, while the TTI kernel is nearly compute bound
(see rooflines in \cite{louboutin2016ppf, devito-api, devito-compiler})
making it more computationally efficient, in particular in combination
with MPI. Future work with involve working on the InfiniBand enabled and
true HPC VM on azure to achieve Cloud performance on par with state of
the art Cluster performance. From the performance obtained in this
experiment, and assuming a fairly standard setup of 5000 independent
source experiments, computing a synthetic dataset would require 322
EFLOP to be computed (23k time-steps x 2.8TFLOP/time-step x 500
sources), or using the full scalability and computing all sources in
parallel would lead to 5.5PFLOP/s from our single source experiment
result.

\section{Performance comparison with other
codes}\label{performance-comparison-with-other-codes}

Because earlier performance benchmarks focus mainly on roofline model
analysis, we also compared the runtime of
\href{https://github.com/devitocodes/devito}{Devito} with a reference
open source hand-coded propagator in collaboration with its author for
completeness. This propagator, described in \cite{thorbecke} is a
state of the art elastic kernel (Equation~\ref{elas1}). For a fair
comparison, we ensure, in collaboration with the author of
\href{https://github.com/JanThorbecke/OpenSource.git}{fdelmodc} that the
physical and computational settings were identical. The setting were the
following:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  2001 by 1001 physical grid points.
\item
  200 grid points of dampening layer (absorbing layer \cite{cerjan}) on
  all four sides (total of 2401x1401 computational grid points).
\item
  10001 time steps.
\item
  Single point source, 2001 receivers.
\item
  Same compiler (GCC/ICC) to compile
  \href{https://github.com/JanThorbecke/OpenSource.git}{fdelmodc} and
  run Devito.
\item
  Intel(R) Xeon(R) CPU E3-1270 v6 @ 3.8GHz.
\item
  Single socket, four physical cores, four physical threads, thread
  pinning to cores and hyperthreading off.
\end{itemize}

The runtime obtained for this problem for both propagators were
identical with less than one percent of difference. This similar runtime
were obtained both with the Intel compiler and the GNU compiler and we
ran the experiment with a fourth order and a sixth order discretization.
The kernels were executed five time each to ensure consistency between
the results and we consistently obtained similar runtimes. This
comparison illustrates the performance achieved with
\href{https://github.com/devitocodes/devito}{Devito} is at least on par
with hand-coded propagators. Considering we do not take adantage of the
\href{https://github.com/devitocodes/devito}{Devito} compiler full
capabilities in the two dimensional case, we are confident that the
generated code will be at least on par with hand-coded version for three
dimensional probalems and this comparison will be part of future work.

\section{Conclusions}\label{conclusions}

The transition from academic size problems, such as the two-dimensional
acoustic wave-equation, to real-world application can be challenging in
particular when attempted as an afterthought. In this work, we showed
that thanks to design principles that aimed at this type of application
from the beginning, \href{https://github.com/devitocodes/devito}{Devito}
provides the scalability necessary. First, we showed that we provide a
high-level interface not only for simple scalar equations but for
coupled PDEs and allow the expression of non-trivial differential
operators in a simple and concise way. Second and most importantly, we
demonstrated that the compiler enables large-scale with state-of-the art
computational performance and programming paradigm. The single-node
performance is on par with hand-coded codes while providing the
flexibility from the symbolic interface, and multi-node parallelism is
integrated in the compiler and interface in a conservative way. Finally,
we demonstrated that our abstraction provides to performance portability
that enables both on-premise HPC and Cloud HPC.

\bibliography{sc20_paper}

\end{document}